---
title: "Lab 2"
author: "An Bui"
date: "17 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 0. Attach packages.
```{r packages, include = FALSE}
library(tidyverse)
library(janitor)
library(MASS)
library(ggrepel)
library(RColorBrewer)
library(factoextra)
library(ggbiplot)
library(shinythemes)
```

To override (set) a function manually throughout an entire workspace, assign the function name explicitly:
```{r assign a function name manually}
select <- dplyr::select
```

Alternatively, you could include `dplyr::select` everywhere to choose dplyr's `select`, and no other package's.

#### 1. Ordinal logisic regression for political party affiliation by ethnicity, age, sex
```{r political data, include = FALSE}
# read in data
pol_party <- read_csv("pol_party.csv") %>% 
  # in janitor package, converts all column names to snake case
  clean_names() %>% 
  # select participant_id through read_carefully columns
  select(participant_id:read_carefully)
```

How do age, ethnicity, and gender influence the liklihood of identifying somewhere on a political spectrum from very liberal to very conservative?

```{r wrangling, include = FALSE}
#### create a new data frame ####
pol_df <- pol_party %>% 
  select(birth_year, gender, ethnicity, political_party) %>% 
  # filter to only include data with birth_year > 1900
  filter(birth_year > 1900,
         # take out NAs from political_party
         political_party != "NA",
         # only include political party less than or equal to 7
         political_party <= 7) %>% 
  # add new column for age
  mutate(age = 2016 - birth_year)
```






















