---
title: "Lab 2"
author: "An Bui"
date: "17 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### 0. Attach packages.
```{r packages, include = FALSE}
library(tidyverse)
library(janitor)
library(MASS)
library(ggrepel)
library(RColorBrewer)
library(factoextra)
library(ggbiplot)
library(shinythemes)
```

To override (set) a function manually throughout an entire workspace, assign the function name explicitly:
```{r assign a function name manually}
select <- dplyr::select
```

Alternatively, you could include `dplyr::select` everywhere to choose dplyr's `select`, and no other package's.

#### 1. Ordinal logisic regression for political party affiliation by ethnicity, age, sex
```{r political data, include = FALSE}
# read in data
pol_party <- read_csv("pol_party.csv") %>% 
  # in janitor package, converts all column names to snake case
  clean_names() %>% 
  # select participant_id through read_carefully columns
  select(participant_id:read_carefully)
```

How do age, ethnicity, and gender influence the liklihood of identifying somewhere on a political spectrum from very liberal to very conservative?  

**Political party**: 1 = very liberal, 7 = very conservative

```{r wrangling, include = FALSE}
#### create a new data frame ####
pol_df <- pol_party %>% 
  select(birth_year, gender, ethnicity, political_party) %>% 
  # filter to only include data with birth_year > 1900
  filter(birth_year > 1900,
         # take out NAs from political_party
         political_party != "NA",
         # only include political party less than or equal to 7
         political_party <= 7) %>% 
  # add new column for age
  mutate(age = 2016 - birth_year)

# When doing OLR, make the dependent variable class ORDERED FACTOR

#### make dependent variable an ordered factor ####
# coerces political_party column to factor
pol_df$political_party <- factor(pol_df$political_party, 
                                 # make that factor ordered
                                 ordered = TRUE,
                                 # call the levels
                                 levels = c("1", "2", "3", "4", "5", "6", "7"))
# check your work with class(pol_df$political_party), should come out as an ordered factor

#### update categorical predictor variables to factors ####
pol_df$ethnicity <- factor(pol_df$ethnicity)
pol_df$gender <- factor(pol_df$gender)
```

```{r basic exploration, include = FALSE}
#### ethnicity counts ####
counts_eth <- pol_df %>% 
  group_by(ethnicity, political_party) %>% 
  # tells you counts of each poltical_party levels for each ethnicity
  tally()

# hey, no shit, outcomes are weighted by white responses because there are way more white people who responded

#### gender counts ####
counts_g <- pol_df %>% 
  group_by(gender, political_party) %>% 
  tally()
```

```{r exploratory graphs, echo = FALSE}
#### histogram ####
# name plot, use ggplot, name data, name x axis
pol_hist <- ggplot(pol_df, aes(x = political_party)) +
  # histogram, fill columns based on ethnicity
  geom_histogram(aes(fill = ethnicity), stat = "count") +
  # use brewer function
  scale_fill_brewer(palette = "YlGnBu") +
  # group by ethnicity, adjust scales for sample sizes
  facet_wrap(~ethnicity, scales = "free")
pol_hist
# non normally distributed responses, duh
```

Just to be clear, n for these data is too small for groups.

```{r ordinal logistic regression, include = FALSE}
#### use polr() for OLR ####
pol_model <- polr(political_party ~ age + gender + ethnicity, data = pol_df)
summary(pol_model)
exp(coef(pol_model))

```

Note: The coefficients are in log odds and are with respect to reference levels to factors (default was female and Asian).  
In summary, you can look at coefficient values and interpret log odds. For example:  
- As age increases, we would expect to see a slight increase in the likelihood that a person ranks themselves higher on the scale, but is negligible.  
- As you consider male to female, all things held constant, they are likely to rank themselves higher on the liberal to conservative scale.  
By exponentiating coefficients, you can get information on the increase in odds given variables.

















